---
title: "STA 380 Homework 2_ver.Winnie"
author: "Yuxin Li"
date: "8/15/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Author attribution

```{r}
rm(list=ls())
library(tm) 
library(magrittr)
```

```{r}
readerPlain = function(fname){
				readPlain(elem=list(content=readLines(fname)), 
							id=fname, language='en') }
							
## get a list of author directories
author_dirs = Sys.glob('data/ReutersC50/C50train/*')

file_list = NULL
#loop through all the documents for each authors
for(author in author_dirs) {
	author_name = substring(author, first=21)
	files_to_add = Sys.glob(paste0(author, '/*.txt'))
	file_list = append(file_list, files_to_add)
}
all_docs = lapply(file_list, readerPlain) 

#name all the documents
mynames = file_list %>%
	{ strsplit(., '/', fixed=TRUE) } %>%
	{ lapply(., tail, n=2) } %>%
	{ lapply(., paste0, collapse = '') } %>%
	unlist
authorname = file_list %>%
	{ strsplit(., '/', fixed=TRUE) } %>%
	{ lapply(.,extract,4)} %>%
	unlist
names(all_docs)= mynames
my_documents = Corpus(VectorSource(all_docs))

#preprocess/tokenize the corpus
my_documents = tm_map(my_documents, content_transformer(tolower)) # make everything lowercase
my_documents = tm_map(my_documents, content_transformer(removeNumbers)) # remove numbers
my_documents = tm_map(my_documents, content_transformer(removePunctuation)) # remove punctuation
my_documents = tm_map(my_documents, content_transformer(stripWhitespace)) ## remove excess white-space
my_documents = tm_map(my_documents, content_transformer(removeWords), stopwords("en")) ##remove stopwords

## create a doc-term-matrix
DTM_all_doc = DocumentTermMatrix(my_documents,control = list(weighting = weightTfIdf))
# inspect(DTM_all_doc[1:10,1:20])
DTM_all_doc = removeSparseTerms(DTM_all_doc, 0.95)

X_train_df = as.data.frame(as.matrix(DTM_all_doc))
author_train = factor(authorname)
```


```{r}
#read in test data

## get a list of author directories
author_dirs_test = Sys.glob('data/ReutersC50/C50test/*')

file_list_test = NULL

#loop through all the documents for each authors
for(author in author_dirs_test) {
	author_name = substring(author, first=21)
	files_to_add = Sys.glob(paste0(author, '/*.txt'))
	file_list_test = append(file_list_test, files_to_add)
}
all_docs_test = lapply(file_list_test, readerPlain) 

#name all the documents
mynames_test = file_list_test %>%
	{ strsplit(., '/', fixed=TRUE) } %>%
	{ lapply(., tail, n=2) } %>%
	{ lapply(., paste0, collapse = '') } %>%
	unlist
authorname_test = file_list_test %>%
	{ strsplit(., '/', fixed=TRUE) } %>%
	{ lapply(.,extract,4)} %>%
	unlist
names(all_docs_test)= mynames_test
my_documents_test = Corpus(VectorSource(all_docs_test))
#preprocess/tokenize the corpus
my_documents_test = tm_map(my_documents_test, content_transformer(tolower)) # make everything lowercase
my_documents_test = tm_map(my_documents_test, content_transformer(removeNumbers)) # remove numbers
my_documents_test = tm_map(my_documents_test, content_transformer(removePunctuation)) # remove punctuation
my_documents_test = tm_map(my_documents_test, content_transformer(removeWords), stopwords("en")) ##remove stopwords
my_documents_test = tm_map(my_documents_test, content_transformer(stripWhitespace)) ## remove excess white-space
## create a doc-term-matrix
DTM_all_doc_test = DocumentTermMatrix(my_documents_test,control = list(weighting = weightTfIdf))
# Remove sparse terms
DTM_all_doc_test = removeSparseTerms(DTM_all_doc_test, 0.95)
#put testing data in dataframe form
X_test_df = as.data.frame(as.matrix(DTM_all_doc_test))
author_test = factor(authorname_test)
```

After preparing the data, we would like to fit a naivebayes model and a logistic model to the data. We would use the naivebayes library to fit the first model. Then, we will compare the percentage of correctness of the two models.

```{r}
#Naive Bayes
library(naivebayes)
nB.model = naive_bayes(author_train~.,data=X_train_df)
nB.pred = data.frame(predict(nB.model,X_test_df))
nB.result = cbind(nB.pred,author_test)
nB.result$correct = (nB.result[,1] == nB.result[,2])
mean(nB.result[,3])
#sum(result[,3])/length(author_test)
```
The Naive Bayes function gives us a out-of-sample accuracy of 44.16%. This means that about 44% of the time the naive bayes classfier will be able to attribute the articles to the correct author.
```{r}
#whose articles are difficult to distinguish?
nB.False = nB.result[nB.result[,3]==FALSE,]
nB.False <- table(nB.False[,2])

print('Frequent mistakes:')
print(sort(nB.False,decreasing = TRUE)/50)
```

As we can tell from the table above, each of these authors have 50 articles included in the training set, David Lawder has the highest misclassfy rate by using Naive Bayes classifier, followed by Jane Macartney.

Let's try a different model to compare the result.

```{r message=FALSE}
#Logistic Regression
library(nnet)
gbm.model = gbm(author_train~.,data=X_train_df,distribution='multinomial',cv.fold2)
gbm.pred = data.frame(predict(gbm.model,X_test_df))
gbm.result = cbind(gbm.pred,author_test)
gbm.result$correct = (gbm.result[,1] == gbm.result[,2])
mean(gbm.result[,3])
```
```{r}
require(gbm)
library(dplyr)
gbm.model = gbm.fit(x=X_train_df,y=author_train,distribution = 'multinomial',n.tree= 50)

```
```{r}

```

